Transformation of forecasts for evaluating predictive performance in an epidemiological context

### 2023-01-17 main.tex

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/SIM-mean-state-size.png}
    \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity. We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the data-generating distribution. 
    These expected CRPS values where computed for different predictive means based on 10,000 samples each and are represented by dots. Solid lines show the corresponding approximation of the expected CRPS based on an assumed normal distribution as discussed in section \ref{sec:methods:vst}. Figure \ref{fig:score-approx} shows the quality of the approximation in more detail. 
    The first distribution (red) is a truncated normal distribution with constant variance (we chose $\sigma = 1$ in order to only obtain positive samples). The second (green) is a negative binomial distribution with variance $\theta = 10$ and variance $\sigma^2 = \mu + 0.1\mu^2$. The third (blue) is a
    Poisson distribution with $\sigma^2 = \mu$. To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one. 
    A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance. Expected CRPS was computed on the natural scale (middle), after adding one and applying a log-transformation to the data (middle) and after applying a square-root transformation (right). B: A but with x axis on the log scale.}
    \label{fig:SIM-wis-state-size-mean}
\end{figure}

@@

In practice, one issue with transformations, such as the log transform, is that they are not readily applicable to negative numbers or zero values, which need to be removed or otherwise handled. 
One common approach to deal with zeros is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm \citep{bellegoDealingLogsZeros2022}. This represents a strictly monotonic transformation and therefore preserves the propriety of the resulting score. The choice of the quantity to add does however influences scores and rankings, as measures of relative errors shrink when adding a constant $a$ to the forecast and the observation. We illustrate this in Figure \ref{fig:illustration-effect-log-offset}. As a rule of thumb, if the observation or the forecast is more than 5 times larger than the offset $a$, then deviations from the natural logarithm without offset should be small, and negligible if it is more than 50 times as large as the offset. Choosing a suitable offset $a$ balances two competing concerns: on the one hand, choosing a small $a$ makes sure that the transformation is as close to a natural logarithm as possible and scores can be interpreted as outlined above. On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero and equalise scores across different orders of magnitude, avoiding giving excessive weight to forecasts for small quantities (see Figure \ref{fig:HUB-log-different-offsets}).

@@

Small discrete observed values can pose a problem as it may be difficult to reliably evaluate relative errors. A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms. This can be seen in Figure \ref{fig:HUB-model-comparison-baseline}, where the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm where forecasts contained zero in its prediction intervals.

@@

\begin{figure}[h!]
\centering
\includegraphics[width = 1\textwidth]{output/figures/illustration-effect-log-ranking-crps.png}
\caption{Illustration of the effect of the log-transformation of the ranking for a single forecast. Shown are CRPS (or WIS, respectively) values as a function of the observed value for two forecasters. Model A issues a geometric distribution (a negative binomial distribution with size parameter $\theta = 1$) with mean $\mu = 10$ and variance $\sigma^2 = \mu + \mu^2 = 110$, while Model B issues a Poisson distribution with mean and variance equal to 10. Zeroes in this illustrative example were handled by adding one before applying the natural logarithm.}
\label{fig:illustration-ranking}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-transformation-regression.png}
    \caption{Relationship between median forecasts and scores. Black dots represent WIS values for two-week ahead predictions of the EuroCOVIDhub-ensemble. Shown in red are the regression lines discussed in Section \ref{sec:HUB-regression} shown in Table \ref{tab:HUB-regression}. A: WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against median predicted values. B: Same as A, with scores obtained after applying a log-transformation to the data. C: Same as A, with scores obtained after applying a square-root-transformation to the data.} 
    \label{fig:HUB-regression}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-epinow.png}
    \caption{
    Forecasts and scores for two-week-ahead predictions from the epiforecasts-EpiNow2 model made in Germany. A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale. B, F: Corresponding scores. C, G: Forecasts and observations on the log scale. D, H: Corresponding scores. 
    }
    \label{fig:HUB-model-comparison-epinow}
\end{figure}

>>>>

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/SIM-mean-state-size.png}
    \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity. We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the data-generating distribution. 
    These expected CRPS values where computed for different predictive means based on 10,000 samples each and are represented by dots. Solid lines show the corresponding approximation of the expected CRPS based on an assumed normal distribution as discussed in section \ref{sec:methods:vst}. Figure \ref{fig:score-approx} shows the quality of the approximation in more detail. 
    The first distribution (red) is a truncated normal distribution with constant variance (we chose $\sigma = 1$ in order to only obtain positive samples). The second (green) is a negative binomial distribution with variance $\theta = 10$ and variance $\sigma^2 = \mu + 0.1\mu^2$. The third (blue) is a
    Poisson distribution with $\sigma^2 = \mu$. To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one. 
    A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance. Expected CRPS was computed on the natural scale (left), after applying a square-root transformation (middle), and after adding one and applying a log-transformation to the data (right). B: A but with x axis on the log scale.}
    \label{fig:SIM-wis-state-size-mean}
\end{figure}

@@

In practice, one issue with transformations, such as the log transform, is that they are not readily applicable to negative numbers or zero values, which need to be removed or otherwise handled. 
One common approach to deal with zeros is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm \citep{bellegoDealingLogsZeros2022}. This represents a strictly monotonic transformation and therefore preserves the propriety of the resulting score. The choice of the quantity to add does however influences scores and rankings, as measures of relative errors shrink when adding a constant $a$ to the forecast and the observation. We illustrate this in Figure \ref{fig:illustration-effect-log-offset}. As a rule of thumb, if if $x > 5a$, the difference between $\log{(x + a)}$ and $\log{(x)}$ is small, and it becomes negligible if $x > 50a$. Choosing a suitable offset $a$ balances two competing concerns: on the one hand, choosing a small $a$ makes sure that the transformation is as close to a natural logarithm as possible and scores can be interpreted as outlined above. On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero, avoiding giving excessive weight to forecasts for small quantities (see Figure \ref{fig:HUB-log-different-offsets}).

@@

A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms. This can be seen in Figure \ref{fig:HUB-model-comparison-baseline}, where the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm where forecasts contained zero in its prediction intervals.

@@

\begin{figure}[h!]
\centering
\includegraphics[width = 1\textwidth]{output/figures/illustration-effect-log-ranking-crps.png}
\caption{Illustration of the effect of the log-transformation of the ranking for a single forecast. Shown are CRPS (or WIS, respectively) values as a function of the observed value for two forecasters. Model A issues a geometric distribution (a negative binomial distribution with size parameter $\theta = 1$) with mean $\mu = 10$ and variance $\sigma^2 = \mu + \mu^2 = 110$), while Model B issues a Poisson distribution with mean and variance equal to 10. Zeroes in this illustrative example were handled by adding one before applying the natural logarithm.}
\label{fig:illustration-ranking}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-transformation-regression.png}
    \caption{Relationship between median forecasts and scores. Black dots represent WIS values for two-week ahead predictions of the EuroCOVIDhub-ensemble. Shown in red are the regression lines discussed in Section \ref{sec:HUB-regression} shown in Table \ref{tab:HUB-regression}. A: WIS for two-week-ahead predictions of the EuroCOVIDhub-ensemble against median predicted values. B: Same as A, with scores obtained after applying a square-root-transformation to the data. C: Same as A, with scores obtained after applying a log-transformation to the data.} 
    \label{fig:HUB-regression}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-epinow.png}
    \caption{
    Forecasts and scores for two-week-ahead predictions from the epiforecasts-EpiNow2 model \citep{epinow2} made in Germany. A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale. B, F: Corresponding scores. C, G: Forecasts and observations on the log scale. D, H: Corresponding scores. 
    }
    \label{fig:HUB-model-comparison-epinow}
\end{figure}
